---
title: "Homework 2: R programming - Question 1"
author: "Viviana Alejandra Rodriguez Romero"
date: "September 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

a) Load the built-in dataset **state** into R's memory.

```{r}    
data(state)
```

b) List the objects in your R session. Get help for the **state** object.

The objects in the R session are:    

```{r}    
ls()
?state
```

c) What type of object is **state.x77**?

```{r}    
class(state.x77)
```    

**state.x77** is a matrix object.     

d) Create a new object called **new.state** that is a **data.frame** based on the data stored in **state.x77**.   

```{r}    
new.state<-as.data.frame(state.x77)
```   

e) Attach the **new.state** object into the search path.    

```{r}    
attach(new.state)
```

f) Compute the mean of the **Illiteracy** vector from the **new.state** object, and its residuals from the mean value.

```{r}    
m<-mean(Illiteracy)
new.state$resid<-Illiteracy-m
```

The mean of the **Illiteracy** is `r m`.  The residuals from the mean of **Illiteracy** are storage on **new.state** as **resid** vector.    

g) Print the elements of **Illiteracy** that are greater than the median illiteracy. Print the elements of **Murder** that are greater than its median.     

The elements of **Illiteracy** that are greater than its median are:    

```{r}    
new.state[Illiteracy>median(Illiteracy),]$Illiteracy
```    

The elements of **Murder** that are greater than its median are:    

```{r}    
new.state[Murder>median(Murder),]$Murder
```   

h) Plot **Illiteracy** on the x-axis and Murder on the y-axis. From interacting with the graph, what 3 states have the highest murder rates?

```{r}
plot(Illiteracy, Murder)
q1.h<-identify(Illiteracy, Murder , labels=row.names(new.state)) # identify points 
```     

The states with the highest murder rates are Alabama, Georgia, and Louisiana.    


i) The plot function has an optional argument, type, that controls what type of plot is made (points, lines, etc). In particular, **type = 'n'** plots nothing, but sets up and draws axes and axis labels. Use this to generate a scatterplot of **Illiteracy** and **Murder**, but with the state abbreviations labeling the points, instead of the default plotting character.

```{r}
plot(Illiteracy, Murder, type = 'n')
text(Illiteracy, Murder,labels=state.abb, cex= 0.5)
```  

j) Write a function is.min that returns a logical vector identifying the minimum values. Write another function (call it **mins**) that calls **is.min** and returns the numerical indices of the minimum values. Apply mins to the **Illiteracy** vector.

```{r}
is.min<-function(v){
v==min(v)  
}

mins<-function(v){
  is.min(v)
}
```     
