---
title: "Homework 2: R programming - Question 1"
author: "Viviana Alejandra Rodriguez Romero"
date: "September 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

a) Load the built-in dataset **state** into R's memory.

```{r}    
data(state)
```
<br />   

b) List the objects in your R session. Get help for the **state** object.

The objects in the R session are:    

```{r}    
ls()
?state
```

<br />     

c) What type of object is **state.x77**?

```{r}    
class(state.x77)
```    

**state.x77** is a matrix object.     

<br />     

d) Create a new object called **new.state** that is a **data.frame** based on the data stored in **state.x77**.   

```{r}    
new.state<-as.data.frame(state.x77)
```   

<br />      

e) Attach the **new.state** object into the search path.    

```{r}    
attach(new.state)
```

<br />       

f) Compute the mean of the **Illiteracy** vector from the **new.state** object, and its residuals from the mean value.

```{r}    
m<-mean(Illiteracy)
new.state$resid<-Illiteracy-m
#The residuals are:
new.state$resid
```

The mean of the **Illiteracy** is `r m`.  The residuals from the mean of **Illiteracy** are storage on **new.state** as **resid** vector.    

<br />      

g) Print the elements of **Illiteracy** that are greater than the median illiteracy. Print the elements of **Murder** that are greater than its median.     

The elements of **Illiteracy** that are greater than its median are:    

```{r}    
new.state[Illiteracy>median(Illiteracy),"Illiteracy",drop=FALSE]
```    

The elements of **Murder** that are greater than its median are:    

```{r}    
new.state[Murder>median(Murder),"Murder",drop=FALSE]
```   

<br />       

h) Plot **Illiteracy** on the x-axis and Murder on the y-axis. From interacting with the graph, what 3 states have the highest murder rates?

```{r}
plot(Illiteracy, Murder)
q1.h<-identify(Illiteracy, Murder , labels=row.names(new.state)) # identify points 
```     

The states with the highest murder rates are Alabama, Georgia, and Louisiana.    

<br />     

i) The plot function has an optional argument, type, that controls what type of plot is made (points, lines, etc). In particular, **type = 'n'** plots nothing, but sets up and draws axes and axis labels. Use this to generate a scatterplot of **Illiteracy** and **Murder**, but with the state abbreviations labeling the points, instead of the default plotting character.

```{r}
plot(Illiteracy, Murder, type = 'n')
text(Illiteracy, Murder,labels=state.abb, cex= 0.5)
```  

<br />      

j) Write a function is.min that returns a logical vector identifying the minimum values. Write another function (call it **mins**) that calls **is.min** and returns the numerical indices of the minimum values. Apply mins to the **Illiteracy** vector.

```{r}
is.min<-function(v){
   v==min(v)  
}

mins<-function(d, v){
  #d: dataset
  #v: variable
  a<-which(is.min(v))
  b<-rownames(d[is.min(v),])
  c<-cbind(a,b)
  return(c)
}

mins(d=new.state, v=Illiteracy)
```     

The minimum value of Illiteracy is `r min(Illiteracy)` on the states Iowa, Nevada and South Dakota.           

