---
title: "Homework 2: R programming"
author: "Viviana Alejandra Rodriguez Romero"
date: "September 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3

Write a function is R to solve the following problem: One dollar invested at an annual rate of r percent compounded monthly for m months is worth  

$$\Big(1+\frac{r}{1200}\Big)^m$$

How much will one dollar be worth at the end of 6 years at 6% interest? What about at 17%?

```{r}  
q3<-function(r,m ){ 
  #r: annual rate in percent
  #m: time in months
  (1+(r/1200))^m
}
q3(6, 6*12)
q3(17, 6*12)

```  

One dollar invested at an annual rate of 6% will worth 1.43 dollars at the end of 6 years, or 2.75 dollars if the annual rate is 17%.    

## Question 4   

A monthly loan payment is computed from the principle of the loan, p, the number of years of payments y, and the interest rate in percent i, according to the formula

$$payment=p*\frac{(i/1200)*(1+i/1200)^{12*y}}{(1+i/1200)^{12*y}-1}$$

Write an R function to carry out this computation. What would the monthly payment be for a $150,000 mortgage at 6.5% for 30 years?

```{r}  
q4<-function(p,y,i){
  #p: loan
  #y: years of payments
  #i: interest rate in percent
  p*((i/1200)*((1+(i/1200))^(12*y)))/((1+(i/1200))^((12*y)-1))
}

q4(150000, 30, 6.5)
```  

If the mortgage is \$150,000 at 6.5% interest rate for 30 years, the monthly payment would be \$816.90. 